---
title: 'The Beer Necessities'
author: "Team: Carly Seeck, Drew Hirsh,  Eric Hedgren, Henry Eichten, Gavin Fisher"
date: "11/27/2023"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(readxl)
library(viridis)
library(tidyr)
library(dplyr)

limit <- 100
```

Importing the Dataset.
```{r}
data <- read.csv("./beer_profile_and_ratings.csv")
head(data, 3)
```

Q: Which combination of mouth feel, taste, and flavor/aroma characteristics returns the highest ratings?

```{r}
percents_flavor_data <- data.frame(data["review_overall"])

percents_flavor_data$'%Bitter' <- round(data$Bitter / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Sweet' <- round(data$Sweet / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Sour' <- round(data$Sour / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Salty' <- round(data$Salty / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Fruits' <- round(data$Fruits / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Hoppy' <- round(data$Hoppy / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Spices' <- round(data$Spices / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Malty' <- round(data$Malty / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

limited_percents_flavor_data <- top_n(percents_flavor_data, limit, review_overall)

head(percents_flavor_data)
```

```{r}
flavor_data <- data[order(-data$review_overall), ]
flavor_data <- flavor_data[c("review_overall", "Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]

limited_flavor_data <- top_n(flavor_data, limit, review_overall)

top_10 <- top_n(flavor_data, 10, review_overall)

print(top_10)
```

Data visualized by percent of variable of sum of all variables by raw values.
```{r}
limited_percents_flavor_data <- limited_percents_flavor_data[order(limited_percents_flavor_data$review_overall), ]

limited_percents_flavor_data %>%
    tidyr::gather(variable, value, -review_overall) %>%
    ggplot(aes(x = variable, y = value, fill = variable)) +
        geom_boxplot() +
        theme_minimal()
```

Data visualized by raw values.
```{r}
limited_flavor_data <- limited_flavor_data[order(limited_flavor_data$review_overall), ]

limited_flavor_data %>%
    tidyr::gather(variable, value, -review_overall) %>%
    ggplot(aes(x = variable, y = value, fill = variable)) +
        geom_boxplot() +
        theme_minimal()
```

Q: Can we find what characteristics will likely cause a beer to be rated highly?
```{r}
flavor_model <- lm(review_overall ~ Bitter + Sweet + Sour + Salty + Fruits + Hoppy + Spices + Malty, data = flavor_data)

summary(flavor_model)
```

Predicting based on the models, if an alcohol will have a high 'overall_review'.

First create data of prime flavors. A prime flavor is the best rated beer with 1 column the same then the rest evenly spread.
```{r}
prime_flavors <- data.frame(matrix(ncol = ncol(flavor_data), nrow = 0))
colnames(prime_flavors) <- colnames(flavor_data)
  prime_flavor <- data.frame(matrix(ncol = ncol(flavor_data), nrow = 1))
prime_flavors$Flavor <- character()

create_prime_flavor <- function(column) {
  colnames(prime_flavor) <- colnames(flavor_data)
  best_flavor <- flavor_data[which.max(flavor_data$review_overall),]

  if (tolower(column) == "bitter") {
    prime_flavor['Bitter'] <- best_flavor['Bitter']
  } else {
    prime_flavor['Bitter'] <- mean(flavor_data$'Bitter')
  }
  if (tolower(column) == "sweet") {
    prime_flavor['Sweet'] <- best_flavor['Sweet']
  } else {
    prime_flavor['Sweet'] <- mean(flavor_data$'Sweet')
  }
  if (tolower(column) == "sour") {
    prime_flavor['Sour'] <- best_flavor['Sour']
  } else {
    prime_flavor['Sour'] <- mean(flavor_data$'Sour')
  }
  if (tolower(column) == "salty") {
    prime_flavor['Salty'] <- best_flavor['Salty']
  } else {
    prime_flavor['Salty'] <- mean(flavor_data$'Salty')
  }
  if (tolower(column) == "fruits") {
    prime_flavor['Fruits'] <- best_flavor['Fruits']
  } else {
    prime_flavor['Fruits'] <- mean(flavor_data$'Fruits')
  }
  if (tolower(column) == "hoppy") {
    prime_flavor['Hoppy'] <- best_flavor['Hoppy']
  } else {
    prime_flavor['Hoppy'] <- mean(flavor_data$'Hoppy')
  }
  if (tolower(column) == "spices") {
    prime_flavor['Spices'] <- best_flavor['Spices']
  } else {
    prime_flavor['Spices'] <- mean(flavor_data$'Spices')
  }
  if (tolower(column) == "malty") {
    prime_flavor['Malty'] <- best_flavor['Malty']
  } else {
    prime_flavor['Malty'] <- mean(flavor_data$'Malty')
  }

  return(prime_flavor)
}
```

```{r}
flavors <- c('Bitter', 'Sweet', 'Sour', 'Salty', 'Fruits', 'Hoppy', 'Spices', 'Malty')

for (flavor in flavors) {
  new_prime_flavor <- create_prime_flavor(flavor)
  new_prime_flavor$Flavor <- flavor
  prime_flavors <- rbind(prime_flavors, new_prime_flavor)
}

head(prime_flavors, 8)
```

```{r}
for (row in seq_len(nrow(prime_flavors))) {
  predicted_score <- predict(flavor_model, newdata = prime_flavors[row, ])
  prime_flavors[row, "review_overall"] <- predicted_score
}
head(prime_flavors, 8)
```

```{r}
ggplot(prime_flavors, aes(x = Flavor, y = review_overall, fill = Flavor)) +
        coord_cartesian(ylim=c(3,5)) +
    geom_bar(stat = "identity")
```