---
title: 'The Beer Necessities'
author: "Team: Carly Seeck, Drew Hirsh,  Eric Hedgren, Henry Eichten, Gavin Fisher"
date: "12/12/2023"
output: html_document
---
```{r include=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)
library(caret)
library(readxl)
library(viridis)
library(tidyr)
library(scatterplot3d)
library(class)

data <- read.csv("./beer_profile_and_ratings.csv")
```

Our data set was a beer profile and ratings set which contained 1.5 million reviews for about 3,200 beers. Each beer had numerous categories such as appearance, aroma, palate, taste, and overall review. Besides the characteristic ratings of the beers there was also ABV(Alcohol content of Beer by Volume) and IBU(International Bitterness Unit) scores. We were interested in evaluating this data set due to it being a unique and large set of data which we could ask many questions about.


### How does the Alcohol content of the beer affect its overall rating?

#### Summary Statistics
```{r}
summary(data$ABV)
```

```{r}
#Max ABV beer
max_abv <- max(data$ABV, na.rm = TRUE)
beer_with_max_abv <- data[data$ABV == max_abv, ]
beer_with_max_abv
```


- The minimum ABV in the data set was 0, meaning non-alcoholic beers were present.
- The 1st quartile was 5.0% ABV, meaning that 25% of beers have an ABV of 5.0% or less.
- The median was 6.0% ABV.
- The mean was 6.53% AVV.
- The 3rd quartile was at 7.6% ABV, meaning 75% of the beers have an ABV of 7.6% or less.
- The max ABV was 57.5% from the beer 'Schorschbräu Schorschbock 57%'.


```{r}
cor(data$ABV, data$review_overall, use = "complete.obs")
```

The correlation coefficient between ABV and overall rating is 0.251. This indicates a weak positive relationship between alcohol content and the overall rating. As the ABV increases, there will be a slight tendency for the overall rating to increase as well, but the relationship is not strong.


#### Plots
```{r}
#Histogram for ABV Distribution
hist(data$ABV, main="Distribution of Alcohol By Volume (ABV)", 
     xlab="ABV (%)", ylab="Frequency", col="blue", breaks=30)

# Histogram for Overall Ratings Distribution
hist(data$review_overall, main="Distribution of Overall Ratings", 
     xlab="Overall Rating", ylab="Frequency", col="green", breaks=30)

# Scatter Plot
plot(data$ABV, data$review_overall, main="Scatter Plot of ABV vs Overall Ratings",
     xlab="Alcohol By Volume (ABV) %", ylab="Overall Rating", pch=19, col=rgb(0,0,1,0.4))

```

Interpretation of ABV vs Overall Rating - The scatter plot indicates that there is very little to no evidence to suggest a clear-cut influence of alcohol content on beer ratings. Other characteristics are likely more indicative of the overall ratings of the beer. 

### Which Brewery has the highest rated beers?
```{r echo=FALSE}
highest_rated_by_brewery <- function(name, displayName) {
  result <- data %>% 
  group_by(Brewery) %>% 
  summarize(
   Rating = mean({{name}})
  ) %>%
  arrange(desc(Rating)) %>%
  slice(1:5)
  print(result)
  return(result)
}
```

#### Top 5 Breweries Overall
```{r echo=FALSE}
overall <- highest_rated_by_brewery(review_overall, "Overall")
```

#### Top 5 Breweries by Aroma
```{r echo=FALSE}
aroma <- highest_rated_by_brewery(review_aroma, "Aroma")
```

#### Top 5 Breweries by Appearance
```{r echo=FALSE}
appearance <- highest_rated_by_brewery(review_appearance, "Appearance")
```

#### Top 5 Breweries by Palate
```{r echo=FALSE}
palate <- highest_rated_by_brewery(review_palate, "Palate")
```

#### Top 5 Breweries by Taste
```{r echo=FALSE}
taste <- highest_rated_by_brewery(review_taste, "Taste")
```


While the overall rating gives us a general overview, viewing the frequency of top five rankings by brewery can produce even more valuable insights into the overall quality of a given brewery

```{r echo=FALSE}
df_list <- list(overall, aroma, appearance, palate, taste)
df_concat <- Reduce(
  function(x, y) 
  merge(x, y, all=TRUE),
  df_list)
df_concat %>% 
  group_by(Brewery) %>%
  summarize(
    Count = n()
  ) %>%
  arrange(Count) %>% 
    mutate(Brewery = factor(Brewery, levels = unique(Brewery))) %>%
    ggplot(aes(x=Brewery, y=Count)) +
    geom_bar(stat="identity") +
    coord_flip() +
    theme(legend.position = "none") +
    labs(title = "Top 5 Ranking: Frequency by Brewery")
```

We notice how a few breweries in the overall ranking do not score within the top 5 by any other metrics: Helios Distillery Co. and  Brauerei Zehendner GmbH. We also notice the relative dominance of breweries The Alchemist and Brouwerij Westvleteren, which ranked 3rd and 5th on the overall ranking, respectively.

### How does the IBU score affect the overall score?

#### What is IBU?
IBU stands for International Bitterness Unit, and it is a scale used to quantify the bitterness of beer. The IBU scale measures the amount of bittering compounds, primarily iso-alpha acids, derived from hops in the beer. These compounds contribute to the bitterness of the beer.

The IBU scale typically ranges from 0 to 100 or more, with higher values indicating a more bitter taste.

```{r echo=FALSE}
avg_score_bins <- data %>%
  mutate(
    Avg.IBU = (Min.IBU + Max.IBU)/2,
    IBU_bin = floor(Avg.IBU/10) * 10
  ) %>%
  group_by(IBU_bin) %>%
  summarize(
    avg_score = mean(review_overall)
  )
  
avg_score_bins %>%
  ggplot(aes(x = IBU_bin, y = avg_score)) +
  geom_bar(stat = "identity") +
  coord_cartesian(ylim = c(min(avg_score_bins$avg_score), max(avg_score_bins$avg_score))) +
  labs(title = "IBU and Overall Score", x = "IBU", y = "Average Score")
```

A histogram plotting the average rating per IBU bin demonstrates that there exists an "ideal" and "least ideal" IBU range, those being 60-69 and 10-19, respectively. The graph also indicates a positive correlation between IBU and average score, in general.

### Can we predict the rating of a beer based on its characteristics?

#### Linear Regression Model
```{r}
features <- data[, c("Sweet", "Bitter", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty", "review_overall")]

# Split data into training and testing sets
set.seed(123)
training_rows <- createDataPartition(features$review_overall, p = 0.8, list = FALSE)
train_data <- features[training_rows, ]
test_data <- features[-training_rows, ]

# Linear Model
model <- lm(review_overall ~ Sweet + Bitter + Sour + Salty + Fruits + Hoppy + Spices + Malty, data = train_data)

summary(model)

# Make predictions on the test set
predictions <- predict(model, test_data)

# Calculate Root Mean Squared Error
rmse <- sqrt(mean((predictions - test_data$review_overall)^2))
print(paste("Root Mean Squared Error:", rmse))
```

### Interpretation of Linear Regression Statistics:
All of the categories except salty had an extremely low p-value (<0.05) which means that 7 out of the 8 characteristics are statistical significant when predicting the overall review. I am going to rank them by significance and leave out salty due to the p-value.

- 1. Bitter (0.0043145): Bitter had the highest positive influence on the overall review. This indicates that consumers generally lean towards the bitter flavor in beer.

- 2. Fruits (0.0035742): Fruitiness is also a significant predictor. Beers that have a fruity character are generally reviewed more favorably.

- 3. Malty (0.0023470): This characteristic has a positive influence on the overall review. Beers with higher malty flavors are more likely to be rated higher, suggesting a strong preference for maltiness.

- 4. Spices (0.0020623): Spiciness is the fourth most influential factor. A greater presence of spices in beer is associated with higher reviews.

- 5. Sour (0.0017782): Sourness has a positive influence on beer reviews. This suggests that beers with higher levels of sourness tend to be rated slightly more favorably.

- 6. Hoppy (-0.0011367): Hoppiness has a slightly negative influence. Beers with pronounced hop flavors tend to receive poorer reviews.

- 7. Sweet (-0.0011201): Sweetness shows a negative impact on ratings. Beers that are perceived as too sweet tend to be rated lower, indicating a possible aversion to excessive sweetness.


I have decided to dive deeper into three of the categories to get a better understanding of how they affect beer ratings. I am choosing 'Bitter', 'Fruits', and 'Malty' as my characteristics to investigate further because they have the highest affect on overall ratings.

### Bitter

#### Summary Statistics
```{r}
bitter_stats <- summary(data$Bitter)
bitter_stats
```

#### Plots
```{r}
ggplot(data, aes(x = Bitter, y = review_overall)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "green") +
  labs(title = "Relationship Between Bitter Scores and Overall Ratings",
       x = "Bitter Score",
       y = "Overall Rating")
```

#### Predicting Example
```{r}
# Linear regression with bitter as the predictor
bitter_regression <- lm(review_overall ~ Bitter, data = data)

# Predicting the rating for a beer with a bitter score of 120 (high)
predicted_rating_bitter <- predict(bitter_regression, newdata = data.frame(Bitter = 120))
predicted_rating_bitter

# Predicting the rating for a beer with a bitter score of 3 (low)
predicted_rating_bitter_2 <- predict(bitter_regression, newdata = data.frame(Bitter = 3))
predicted_rating_bitter_2

```

### Fruits

#### Summary Statistis
```{r}
fruits_stats <- summary(data$Fruits)
fruits_stats
```

#### Plots
```{r}
ggplot(data, aes(x = Fruits, y = review_overall)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Relationship Between Fruit Scores and Overall Ratings",
       x = "Fruit Score",
       y = "Overall Rating")

```

#### Predicting Example
```{r}
# Linear regression with fruit as the predictor
fruits_regression <- lm(review_overall ~ Fruits, data = data)

# Predicting the rating for a beer with a fruit score of 110 (high)
predicted_rating_fruits <- predict(fruits_regression, newdata = data.frame(Fruits = 110))
predicted_rating_fruits

# Predicting the rating for a beer with a fruit score of 3 (low)
predicted_rating_fruits_2 <- predict(fruits_regression, newdata = data.frame(Fruits = 3))
predicted_rating_fruits_2

```

### Malty

#### Summary Statistics
```{r}
malty <- summary(data$Malty)
malty
```

#### Plots
```{r}
ggplot(data, aes(x = Malty, y = review_overall)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "blue") +
  labs(title = "Relationship Between Malty Scores and Overall Ratings",
       x = "Malty Score",
       y = "Overall Rating")
```

#### Predicting Example
```{r}
# Linear regression with malty as the predictor
malty_regression <- lm(review_overall ~ Malty, data = data)

# Predicting the rating for a beer with a malty score of 200 (very high)
predicted_rating_malty <- predict(malty_regression, newdata = data.frame(Malty = 200))
predicted_rating_malty

#Predicting the rating for a beer with a malty score of 30 (very low).
predicted_rating_malty_2 <- predict(malty_regression, newdata = data.frame(Malty = 30))
predicted_rating_malty_2
```




### Which beers have the highest aroma, appearance, palate, and taste ratings?
```{r echo=FALSE}
# top ten beers from each category
aromaTop <- data[c("Name", "review_aroma")]
aromaTop <- aromaTop %>% arrange(desc(review_aroma)) %>% slice(1:10)
colnames(aromaTop)[2] <- "Aroma Rating"

appearanceTop <- data[c("Name", "review_appearance")]
appearanceTop <- appearanceTop %>% arrange(desc(review_appearance)) %>% slice(1:10)
colnames(appearanceTop)[2] <- "Appearance Rating"


palateTop <- data[c("Name", "review_palate")]
palateTop <- palateTop %>% arrange(desc(review_palate)) %>% slice(1:10)
colnames(palateTop)[2] <- "Palate Rating"

tasteTop <- data[c("Name", "review_taste")]
tasteTop <- tasteTop %>% arrange(desc(review_taste)) %>% slice(1:10)
colnames(tasteTop)[2] <- "Taste Rating"
```

```{r echo=FALSE}
kable(aromaTop, caption = "Top 10 Aroma Rated Beers", col.names = c("Name", "Aroma Rating"), format = "html", border = 2) %>% kable_styling()
```

```{r echo=FALSE}
kable(appearanceTop, caption = "Top 10 Appearance Rated Beers", col.names = c("Name", "Appearance Rating"), format = "html", border = 2) %>% kable_styling()
```

```{r echo=FALSE}
kable(palateTop, caption = "Top 10 Palate Rated Beers", col.names = c("Name", "Palate Rating"), format = "html", border = 2) %>% kable_styling()
```

```{r echo=FALSE}
kable(tasteTop, caption = "Top 10 Taste Rated Beers", col.names = c("Name", "Taste Rating"), format = "html", border = 2) %>% kable_styling()
```

### What are the most common characteristics of beer?

The following characteristics are defined by word counts found in up to 25 reviews of each beer. The assumption is that people writing reviews are more than likely describing what they do experience rather than what they do not. (Refer to the file Beer Descriptors Simplified to see the list of words that are used to calculate the values contained in each of the feature columns below)

(Mouthfeel)
9 - Astringency
10 - Body
11 - Alcohol

(Taste)
12 - Bitter
13 - Sweet
14 - Sour
15 - Salty

(Flavor And Aroma)
16 - Fruits
17 - Hoppy
18 - Spices
19 - Malty

```{r echo=FALSE}
characteristicsRank <- colSums(data[,c("Astringency", "Body", "Alcohol", "Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")])

characteristicsRank <- sort(characteristicsRank, decreasing = TRUE)

characteristicsRank <- stack(characteristicsRank)

colnames(characteristicsRank)[1] <- "Frequency"
colnames(characteristicsRank)[2] <- "Characteristic"
characteristicsRank <- characteristicsRank %>% select(Characteristic, Frequency)

kable(characteristicsRank, caption = "Characteristic Frequencies", col.names = c("Characteristic", "Frequency"), format = "html", border = 2) %>% kable_styling()
```



Q: Which combination of mouth feel, taste, and flavor/aroma characteristics returns the highest ratings?

```{r echo=FALSE}
percents_flavor_data <- data.frame(data["review_overall"])

percents_flavor_data$'%Bitter' <- round(data$Bitter / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Sweet' <- round(data$Sweet / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Sour' <- round(data$Sour / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Salty' <- round(data$Salty / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Fruits' <- round(data$Fruits / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Hoppy' <- round(data$Hoppy / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Spices' <- round(data$Spices / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

percents_flavor_data$'%Malty' <- round(data$Malty / rowSums(data[, c("Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]), 5)

limit <- 100
limited_percents_flavor_data <- top_n(percents_flavor_data, limit, review_overall)

head(percents_flavor_data)
```

```{r echo=FALSE}
flavor_data <- data[order(-data$review_overall), ]
flavor_data <- flavor_data[c("review_overall", "Bitter", "Sweet", "Sour", "Salty", "Fruits", "Hoppy", "Spices", "Malty")]

limited_flavor_data <- top_n(flavor_data, limit, review_overall)

top_10 <- top_n(flavor_data, 10, review_overall)

print(top_10)
```

Data visualized by percent of variable of sum of all variables by raw values.

Reading the boxes as heat maps it can be understood how the score of a drinks overall_review can be predicted with mixtures of flavors.

This shows significance in %'s.
```{r echo=TRUE}
limited_percents_flavor_data <- limited_percents_flavor_data[order(limited_percents_flavor_data$review_overall), ]

limited_percents_flavor_data %>%
    tidyr::gather(variable, value, -review_overall) %>%
    ggplot(aes(x = variable, y = value, fill = variable)) +
        geom_boxplot() +
        theme_minimal()
```

This shows significance in raw amount in drinks.
```{r echo=FALSE}
limited_flavor_data <- limited_flavor_data[order(limited_flavor_data$review_overall), ]

limited_flavor_data %>%
    tidyr::gather(variable, value, -review_overall) %>%
    ggplot(aes(x = variable, y = value, fill = variable)) +
        geom_boxplot() +
        theme_minimal()
```

Q: Can we find what characteristics will likely cause a beer to be rated highly?


We will determine flavor significance by creating perfect drinks with specific values being of the best.

First a model must be created.
```{r echo=TRUE}
flavor_model <- lm(review_overall ~ Bitter + Sweet + Sour + Salty + Fruits + Hoppy + Spices + Malty, data = flavor_data)

summary(flavor_model)
```

```{r echo=FALSE}
prime_flavors <- data.frame(matrix(ncol = ncol(flavor_data), nrow = 0))
colnames(prime_flavors) <- colnames(flavor_data)
  prime_flavor <- data.frame(matrix(ncol = ncol(flavor_data), nrow = 1))
prime_flavors$Flavor <- character()

create_prime_flavor <- function(column) {
  colnames(prime_flavor) <- colnames(flavor_data)
  best_flavor <- flavor_data[which.max(flavor_data$review_overall),]

  if (tolower(column) == "bitter") {
    prime_flavor['Bitter'] <- best_flavor['Bitter']
  } else {
    prime_flavor['Bitter'] <- mean(flavor_data$'Bitter')
  }
  if (tolower(column) == "sweet") {
    prime_flavor['Sweet'] <- best_flavor['Sweet']
  } else {
    prime_flavor['Sweet'] <- mean(flavor_data$'Sweet')
  }
  if (tolower(column) == "sour") {
    prime_flavor['Sour'] <- best_flavor['Sour']
  } else {
    prime_flavor['Sour'] <- mean(flavor_data$'Sour')
  }
  if (tolower(column) == "salty") {
    prime_flavor['Salty'] <- best_flavor['Salty']
  } else {
    prime_flavor['Salty'] <- mean(flavor_data$'Salty')
  }
  if (tolower(column) == "fruits") {
    prime_flavor['Fruits'] <- best_flavor['Fruits']
  } else {
    prime_flavor['Fruits'] <- mean(flavor_data$'Fruits')
  }
  if (tolower(column) == "hoppy") {
    prime_flavor['Hoppy'] <- best_flavor['Hoppy']
  } else {
    prime_flavor['Hoppy'] <- mean(flavor_data$'Hoppy')
  }
  if (tolower(column) == "spices") {
    prime_flavor['Spices'] <- best_flavor['Spices']
  } else {
    prime_flavor['Spices'] <- mean(flavor_data$'Spices')
  }
  if (tolower(column) == "malty") {
    prime_flavor['Malty'] <- best_flavor['Malty']
  } else {
    prime_flavor['Malty'] <- mean(flavor_data$'Malty')
  }

  return(prime_flavor)
}
```

Prime drinks are created with a specific flavor being of the best reviewed drink, then the rest of the values are average of all other drinks.

```{r echo=TRUE}
flavors <- c('Bitter', 'Sweet', 'Sour', 'Salty', 'Fruits', 'Hoppy', 'Spices', 'Malty', 'Base')

for (flavor in flavors) {
  new_prime_flavor <- create_prime_flavor(flavor)
  new_prime_flavor$Flavor <- flavor
  prime_flavors <- rbind(prime_flavors, new_prime_flavor)
}

head(prime_flavors, 9)
```

```{r echo=FALSE}
for (row in seq_len(nrow(prime_flavors))) {
  predicted_score <- predict(flavor_model, newdata = prime_flavors[row, ])
  prime_flavors[row, "review_overall"] <- predicted_score
}
head(prime_flavors, 8)
```

Flavor significance can be understood by height difference in review scores of better than average flavors being the best ingredient.
```{r echo=FALSE}
ggplot(prime_flavors, aes(x = Flavor, y = review_overall, fill = Flavor)) +
        coord_cartesian(ylim=c(3,4)) +
    geom_bar(stat = "identity") +
  theme(legend.position = "none")
```

What are the most popular beer styles?

To find the most popular beer styles, we simply looked at the "style" category in the data set which is the style of beer. Below is the highest rated style and the lowest rated style. We then graphed the top 15 styles and bottom 15 styles.

```{r echo=FALSE}
data <- read.csv("beer_profile_and_ratings.csv")

beer_ratings <- data %>%
  group_by(Style) %>%
  summarise(Avg_Review = mean(review_overall, na.rm = TRUE))

highest_rated_style <- beer_ratings[which.max(beer_ratings$Avg_Review), ]
lowest_rated_style <- beer_ratings[which.min(beer_ratings$Avg_Review), ]
print(highest_rated_style)
print(lowest_rated_style)
```

Graph for 15 highest rated styles
```{r echo=FALSE}
# graph for 15 highest rated styles
top_15_styles <- beer_ratings %>%
  arrange(desc(Avg_Review)) %>%
  head(15)
print(top_15_styles)
ggplot(top_15_styles, aes(x=reorder(Style, Avg_Review), y=Avg_Review, fill=Style)) +
  geom_bar(stat="identity", fill="gray") +
  coord_flip() +
  labs(x="Beer Style", y="Average Review", title="Top 15 Beer Styles by Average Review") +
  theme_minimal() +
  theme(legend.position = "none")
```

Graph for 15 lowest rated styles
```{r echo=FALSE}
# graph for 15 lowest rated styles
bottom_15_styles <- beer_ratings %>%
  arrange(Avg_Review) %>%
  head(15)
print(bottom_15_styles)
ggplot(bottom_15_styles, aes(x=reorder(Style, -Avg_Review), y=Avg_Review, fill=Style)) +
  geom_bar(stat="identity", fill="grey") +
  coord_flip() +
  labs(x="Beer Style", y="Average Review", title="Bottom 15 Beer Styles by Average Review") +
  theme_minimal() +
  theme(legend.position = "none")
```


Can we recommend other beers if someone picks a beer they like?

This question was answered with calculating euclidean distance between scores of different variables in the data set. For example, below we return the 5 most similar beers to Long Trail Ale based on sweet and sour scores.

```{r echo=FALSE}
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

selected_columns <- c('Sweet', 'Sour')

beer_data_norm <- as.data.frame(lapply(data[, selected_columns], normalize))

user_beer_row <- which(data$`Name` == "Long Trail Ale")

if(length(user_beer_row) == 0) {
  print("The beer 'Long Trail Ale' does not exist in the data.")
} else {
  user_beer <- beer_data_norm[user_beer_row, ]

  beer_data_norm <- beer_data_norm[-user_beer_row, ]

  dist_matrix <- as.matrix(dist(rbind(user_beer, beer_data_norm), method = "euclidean"))
  dist_vector <- as.vector(dist_matrix[1, -1])
  closest_beers <- order(dist_vector)[1:5]
  
  # print(dist_vector[closest_beers])
  print(beer_data_norm[closest_beers, ])
  # print(user_beer)

  
  recommended_beers <- data[closest_beers, ]
  print(recommended_beers)
}
```

The graph below shows sweet and sour similarity toward Long Trail Ale.

```{r echo=FALSE}
data <- read.csv("beer_profile_and_ratings.csv")
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
selected_columns <- c('Sweet', 'Sour')
beer_data_norm <- as.data.frame(lapply(data[, selected_columns], normalize))
user_beer_row <- which(data$`Name` == "Long Trail Ale")
user_beer <- beer_data_norm[user_beer_row, ]

dist_matrix <- as.matrix(dist(rbind(user_beer, beer_data_norm), method = "euclidean"))
dist_vector <- as.vector(dist_matrix[1, -1])
beer_data_norm$Distance <- c(dist_vector[1:(user_beer_row-2)], NA, dist_vector[user_beer_row:(length(dist_vector))])

ggplot(beer_data_norm, aes(x=Sweet, y=Sour, color=Distance)) +
  geom_point() +
  scale_color_gradient(low = "blue", high = "red") +
  labs(x="Sweet", y="Sour", color="Distance to Long Trail Ale") +
  theme_minimal()

```

The graph below is a 3d representation of three characteristics, sweet, sour and bitter vs Long Trail Ale.

```{r echo=FALSE}
data <- read.csv("beer_profile_and_ratings.csv")

# Normalize the data
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}
selected_columns <- c('Sweet', 'Sour', 'Bitter')
beer_data_norm <- as.data.frame(lapply(data[, selected_columns], normalize))

# Select a beer
user_beer_row <- which(data$`Name` == "Long Trail Ale")
user_beer <- beer_data_norm[user_beer_row, ]

# Remove the user's beer from the data
beer_data_norm <- beer_data_norm[-user_beer_row, ]

# Calculate the Euclidean distances
dist_matrix <- as.matrix(dist(rbind(user_beer, beer_data_norm), method = "euclidean"))
dist_vector <- as.vector(dist_matrix[1, -1])

# Add the distances to the data frame
beer_data_norm$Distance <- c(dist_vector[1:(user_beer_row-1)], dist_vector[user_beer_row:length(dist_vector)])

# Create a 3D scatterplot
scatterplot3d(beer_data_norm$Sweet, beer_data_norm$Sour, beer_data_norm$Bitter, color=rainbow(length(beer_data_norm$Distance))[rank(beer_data_norm$Distance)], xlab="Sweet", ylab="Sour", zlab="Bitter")
```

Finally we find the 5 closest beers to Long Trail Ale using every characteristic in the dataset.

```{r echo=FALSE}
normalize <- function(x) {
  return ((x - min(x)) / (max(x) - min(x)))
}

selected_columns <- c('Bitter','Sweet','Sour','Salty','Fruits','Hoppy','Spices','Malty')

beer_data_norm <- as.data.frame(lapply(data[, selected_columns], normalize))

user_beer_row <- which(data$`Name` == "Long Trail Ale")

if(length(user_beer_row) == 0) {
  print("The beer 'Long Trail Ale' does not exist in the data.")
} else {
  user_beer <- beer_data_norm[user_beer_row, ]

  beer_data_norm <- beer_data_norm[-user_beer_row, ]

  dist_matrix <- as.matrix(dist(rbind(user_beer, beer_data_norm), method = "euclidean"))
  dist_vector <- as.vector(dist_matrix[1, -1])
  closest_beers <- order(dist_vector)[1:5]
  
  result <- data.frame(Name = c(data[user_beer_row, "Name"], data[closest_beers, "Name"]),
                       beer_data_norm[c(user_beer_row, closest_beers), ],
                       Distance = c(NA, dist_vector[closest_beers]))
  
  print(result)
}
```

